Palomar 5 Gaps

### 2024-10-22 paper/main.tex
cleaned the discussion on the perturber stats

    We desire to search for some generalizable characterisitcs of the perturbers to see what contrasts them to globular clusters that do not impact the Pal5 Stream. We present a suite of figures to answer for this end in Fig.~\ref{fig:mass_size_plane}.

@@

      From the top left panel of Fig.~\ref{fig:mass_size_plane}, the immediate result is that all clusters that impact Palomar~5's stream are within a band of orbital energies from about $-2~\textrm{ to }~-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$. This is non-surprising result indicates that in order for a cluster to perturb Palomar 5's stream, it must enter Palomar~5's orbital space. Additionally, we reproduce the results of \citet{2021MNRAS.505.5978V}, who found a wide range of possible orbital solutions for the least bound clusters, some being retrograde and some prograde. For this potential model, only 3 samplings of a globular cluster had positive energies.

      From the top left panel of Fig.~\ref{fig:mass_size_plane} shows the mass-size plane of the perturbers against those that do not provoke gaps. We find a lower limit of under $10^5~\textrm{M}_{\odot}$ for provoking gaps. Mass and size are not the full picture, since provoking a gap is given by how much momentum is injected into the stream by Eq.~\ref{eq:change_in_momentum}. However, this emperical lower limit is a practical result since the relative velocity and impact parameter are known values since they happen so long ago in the past. The bottom left panel shows a result which is more true to the nature of the physics of the problem. By tracking the closest approaches and inspecting their approximate imparted momentum, we find that the gap-creating impacts have systematically higher imparted momenta that those that do not create gaps. However, there is significant overlap. Of course, the three parameter law from ~\ref{eq:change_in_momentum} is a simplified image. \citet{2015MNRAS.450.1136E} extended this to include the characterisitc size of the perturber and separated the relative velocity into perpendicular and parallel components to the stream an subsequently quantified how the parameters effect gap evolution. However, when studying the impact geometry at this level we found no satisfying combination of the parameters that clearly discriminated between the gap and no gap case. An extension this is shown in Fig.~\ref{fig:impact_geometry_statistics}, the only additional insights being that there is an observed maximum impact parameter which is about ~100~parsecs. This is roughly on the order of magnitude of the stream width. Additionally, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of the action-frequencies of the star-particles being impacted plays an important factor. Stream regions that are well separated by orbital frequency accept gaps easier while the regions of the stream that have a wide range of frequencies (those closests to the cluster) erase the history of the impact easier. In essence, an additional criteria is that the imparted change in frequency must be greater than the range in frequencies present in the stream area being hit.

      From the bottom right panel we observe where the impacts happen in the meridional plane lain over Palomar 5's orbit. We see that there is a trend with $r$, the 3D galactocentric radius. The number of counts follows this equation: $N = -2.5 r + 50$ and has a pearson coefficient of -0.86, so quite strong I must say.

>>>>

    We aim to identify general characteristics of perturbers that distinguish them from globular clusters that do not affect the Palomar 5 stream. Fig.~\ref{fig:mass_size_plane} presents several analyses to explore this.

@@

      The top-left panel shows that all clusters impacting the stream fall within a specific range of orbital energies, between $-2$ and $-1 \times10^5~\textrm{km}^2\textrm{s}^{-2}$. This is expected, as only clusters within Palomar 5's orbital space can interact with the stream. Our results are consistent with the findings from \citet{2021MNRAS.505.5978V}, who identified a wide range of orbital solutions for the least bound clusters, including both retrograde and prograde orbits. In our potential model, only three clusters samplings had positive orbital energies.

      The mass-size relation, shown in the top-right panel, reveals a lower limit of $10^5~\textrm{M}_{\odot}$ for clusters that provoke gaps. While mass and size alone are insufficient predictors, this empirical limit is useful because relative velocity and impact parameters are historical values. The bottom-left panel offers a more physically meaningful result: gap-creating impacts tend to impart more momentum into the stream than non-gap impacts, though with some overlap. While \citet{2015MNRAS.450.1136E} extended the three-parameter model (Eq.~\ref{eq:change_in_momentum}) to include the size of the perturber and the perpendicular and parallel components of relative velocity, we found no clear combination of these factors that distinguishes gap from non-gap cases.

      Further insights are shown in Fig.\ref{fig:impact_geometry_statistics}, highlighting a maximum impact parameter of around 200~pc, roughly the width of the stream. Additionally, as \citet{2016MNRAS.457.3817S} demonstrated, the dispersion of action-frequencies in the stream plays a role. Regions of the stream with well-separated frequencies are more susceptible to gap formation, while those with a wide frequency range (near the cluster) tend to erase the history of impacts. Thus, for a gap to form, the imparted change in frequency must exceed the range of frequencies in the impacted region. Thus, the strong flybys that occured close to the cluster and were thus inconsequential for gap formation.

      The bottom-right panel reveals a strong negative correlation between the location of impacts and the galactocentric radius, $r$. The number of gap creating impacts follows the relation $N = -2.5r + 50$, with a Pearson coefficient of -0.86. This is a non-obvious result. While there are more clusters near the galactic center, orbits naturally spend more time near their apocenters, and the lower relative velocities in these regions should favor gap creation. However, this result suggests that cluster galactic number density outweighs these factors in determining the number of gaps.
